---
title: "Data Exploration"
author: "Fausto Lopez"
date: "April 28, 2017"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(data.table)
library(lubridate)
library(ggthemes)
library(fasttime)
library(bit64)
library(zoo)
options(scipen = 999)


setwd("C:\\Users\\lopezf\\Desktop\\regression_test\\regression_test")
master = fread("master_raw.csv")[,-1][
  ,timestampz:=fastPOSIXct(timestampz, tz = "GMT")][
    ,subway_entries:=as.numeric(subway_entries)][
      ,subway_exits:=as.numeric(subway_exits)]


# 
# [
#   ,zone.f:=as.factor(taxi_zone)][
#     ,hour.n:=as.numeric(as.character(hour))][
#       ,weekday.f:=as.factor(weekday)]
gc()
str(master)
summary(master)
```


##dealing with the missing data

```{r missing_data}

#treat missing values in data--------------------------------------------------------------
#we assume taxi data as pure so NA's for trips = 0
master[,trips_shl:= ifelse(is.na(trips_shl), 0, trips_shl)]
master[,trips_med:= ifelse(is.na(trips_med), 0, trips_med)]
master[,bike_trips:= ifelse(is.na(bike_trips),0,bike_trips)]
master[,total_trips:= ifelse(is.na(total_trips),0,total_trips)]

#code fs variables as 0 if not present 
master[,arts_entertainment:= ifelse(is.na(arts_entertainment), 0, arts_entertainment)]
master[,college_university := ifelse(is.na(college_university), 0 , college_university)]
master[,event:= ifelse(is.na(event), 0, event)]
master[,food := ifelse(is.na(food), 0, food)]
master[,nightlife:= ifelse(is.na(nightlife), 0, nightlife)]
master[,outdoors_rec := ifelse(is.na(outdoors_rec), 0, outdoors_rec)]
master[,professional := ifelse(is.na(professional), 0, professional)]
master[,residence := ifelse(is.na(residence), 0, residence)]
master[,shop_service := ifelse(is.na(shop_service),0, shop_service)]
master[,travel := ifelse(is.na(travel), 0 , travel)]

#total_trips, weekday and zone as factor
master[,weekday.f := as.factor(weekdays(timestampz))][
,zone.f := as.factor(taxi_zone)][
,hour.f :=as.factor(hour(timestampz))][
,hour.n:=as.numeric(hour(timestampz))]


#fill in NA for daily values 
master[,cpi:=na.locf(cpi,fromLast = F)][
  ,sp500:=na.locf(sp500,fromLast = F)][
  ,mean_wind_speed:=na.locf(mean_wind_speed, fromLast = F)][
    ,mean_temp:=na.locf(mean_temp, fromLast = F)][
      ,rain_inches:=na.locf(rain_inches, fromLast = F)][
      ,gas_price:= na.locf(gas_price,fromLast = F)]

gc()


```

## Starting with Trends

First we look at the basic trips per day for each type of service:

```{r trends, echo = F}

#look at trips per day 
daily_volume = master[,.(shl_volume = sum(trips_shl)
          ,yellow_volume = sum(trips_med)), by = .(dayz=substr(timestampz, 1, 10))][
            ,total_volume:=shl_volume + yellow_volume]


ggplot(daily_volume, aes(x = dayz, y = total_volume, group = 1)) + 
 geom_line(aes(y = shl_volume, colour = "shl_trips")) +
  geom_line(aes(y = yellow_volume, colour = "yellow_trips")) +
  geom_line(aes(y = total_volume, colour = "total_trips")) + 
  ggtitle("Trips Per Day")+
theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_minimal()

#trips per week
ggplot(master[,.(shl_volume = sum(trips_shl)
          ,yellow_volume = sum(trips_med))
          ,by = .(week = paste0(
            year(timestampz), "-", week(timestampz)
            )
                  )][
            ,total_volume:=shl_volume + yellow_volume], aes(x = week , y = total_volume, group = 1)) + 
 geom_line(aes(y = shl_volume, colour = "shl_trips")) +
  geom_line(aes(y = yellow_volume, colour = "yellow_trips")) +
  geom_line(aes(y = total_volume, colour = "total_trips")) + 
  ggtitle("Trips Per Week")+
    theme(axis.text.x  = element_text(
      vjust = .5,
      size = 8,
      angle = 90
    )) +
  theme_minimal()


```

Lets break down the trips by hour:

```{r, echo = F}
hourly_volume = master[,.(shl_volume = sum(trips_shl)
          ,yellow_volume = sum(trips_med)), by = .(hour=substr(timestampz, 12, 13))][
            ,total_volume:=shl_volume + yellow_volume]

ggplot(hourly_volume, aes(x = hour, y = total_volume, group = 1)) + 
 geom_line(aes(y = shl_volume, colour = "shl_trips")) +
  geom_line(aes(y = yellow_volume, colour = "yellow_trips")) +
  geom_line(aes(y = total_volume, colour = "total_trips")) + 
  ggtitle("Trips Per Hour")+
  theme_minimal()



```


We can combine these two to create an animation that plays out what pickups look like over time by location:

```{r}

```


How have business density changed over time?

```{r}

```

How has citibike usage changed over time?



##Understanding underserved zones





##Bivariate Analysis and Distributions

```{r, distributions, echo = F}
par(mfrow=c(3,2))
hist(master$mean_temp, main="temperature")
hist(master$mean_wind_speed, main="wind speed")
hist(master$rain_inches, main="precipitation")
hist(master$gas_price, main = "wholesale gas prices")
hist(master$bike_trips, main = "daily citibike trips")
hist(master$total_trips, main = "daily total trips")

```


```{r}
plot(master[,.(hour.f,gas_price, mean_temp, mean_wind_speed, rain_inches, total_trips)])
gc()
```



```{r}

#zone hour of day combined
master[,zh_c.f:=as.factor(paste0(as.character(zone.f),".",as.character(hour.f)))]

#zone hour of day and weeday combined
master[,zhwk_c.f:=as.factor(paste0(as.character(zone.f),"."
                                   ,as.character(weekday.f),".",as.character(hour.f)))]




```






##Running the basic model on Manhattan


Here is the basic feature set:
```{r basic_model}

master_man = master[boro == "Manhattan",]

#break training and test sets
master_man = master_man[sample(1:nrow(master_man), 100000),]
setorder(master_man, timestampz)
samp = sample(nrow(master_man), 0.6 * nrow(master_man))
train = master_man[samp, ]
test = master_man[-samp, ]

#features: zone, hourofday, dayofweek, daily rain, temperature
fit = lm(total_trips ~  + 
           mean_temp +
           zone.f +
           hour.n +
           weekday.f +
           rain_inches 
           #+ gas_price 
           , train)

summary(fit)

```


We add to the linear regression the combination variables:
```{r}
gc()
fit = lm(total_trips ~  + 
           mean_temp +
           zone.f +
           hour.n +
           weekday.f +
           rain_inches 
           +zh_c.f 
           +zhwk_c.f
           #+ gas_price 
           , train)

summary(fit)

```



##adding in other variables for Manhattan

```{r, echo = F}



fit = lm(total_trips ~  + 
           mean_temp +
           zone.f +
           hour.n +
           weekday.f +
          zh_c.f +
           zhwk_c.f + 
           bike_trips +
           #mean_wind_speed + 
           rain_inches +
           gas_price 
         #+ authorized_vehicles
         + arts_entertainment
         + college_university
         + event
         + food
         + nightlife
         + outdoors_rec
         + professional      
         + residence
         + shop_service
         + travel
         , train)

summary(fit)


```




## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
